Topological Equivariant Artist Model for Visualization Library Architecture

### 2021-11-10 paper/main.tex
new category E that is sets of gamma

Formally, a fiber bundle is a mathematical structure $(\dtotalc, \dbasec, \pi, \dfiberc)$
\begin{equation}
  \label{eq:fiber_bundle}
  \begin{tikzcd}
      \dfiber \arrow[r, hook] & \dtotal \arrow[r, "\pi"] & \dbase
  \end{tikzcd}
\end{equation}

\begin{definition}
The \textcolor{base}{base space} \dbasec\ is a topological space, which means it is a a set $\dbase$ with a collection of open sets surrounding each element in the set. For each point $\dbasepoint in \dbase$, there is a collection of subsets $\{U\} \subset \dbase$ such that $x \in U \subseteq \dbase$ \cite{spanier1989algebraic, bradleyTopologyCategoricalApproach2020}  
\end{definition}
\note{intuition about base space}

@@

How these spaces fit together is illustrated in \autoref{fig:related-work:fiber-bundle}, wherein values lives in the fiber $\dfiber\subseteq\dtotal$. In this example, the fiber \dfiber\ is the cartesian product of two sets $F_{0}\times F_{1}$ where each fiber $F_{i}$ is ....
\begin{equation}
  formal equation of the fiber?
  \label{eq:related-work:fiber}
\end{equation}

\begin{equation}
  \begin{tikzcd}
      \dfiber \arrow[r, hook] & \dtotal \arrow[d, "\pi"'] \\
                        & \dbase \arrow[u, "\dsection"', bend right]
  \end{tikzcd}
\end{equation}

The set of all sections sections $\{\dsection^{i}: \U_{j}\rightarrow \dtotal\}|_{i\in I}$ over an arbitrary open set $U_{j}$ is denoted $\Gamma(U_{\dbasepoint}, \dtotal)$. The set of all sections over the total space \dbase\ is denoted $\Gamma(\dbase, \dtotal)$.

@@

\section{Artist} 
In this section, we use category theory to formally express the implicit assumptions that visualization library components make in transforming data into graphical representations. We propose that the visualization transformation can be modeled as a functor, which we call the \textcolor{artist}{Artist} $\mathcal{\vartist}$\footnote{We call this transformation the artist because the \mintinline{python}{Artist} object in Matplotlib\cite{hunterMatplotlib2DGraphics2007}}. The artist \vartist\ converts data, which we model as the catgory $\mathcal{\dtotal}$ to graphics, which we model as the category $\mathcal{\gtotal}$. We formulate this association as

\begin{equation}
  \vartist: \mathcal{\dtotal} \rightarrow \mathcal{\gtotal}
  \label{eq:artist:artist}
\end{equation}
In this section we describe the structure of the data and graphic spaces, which we formulate as objects in categories and the morphisms between those objects. We then use these definitions to express the structure that an artist must preserve. Finally, we propose that artists can be composed by manufacturing new artists based on various types of input data.

\subsection{Categorical Artist}
\label{sec:artist:category}
We define the categories $\mathcal{\dtotal}$ and $\mathcal{\gtotal}$ as sheaves of sections\cite{SheafMathematics2021,spanier1989algebraic}, of fiberbundles, as introduced in \autoref{sec:related-work:fiber-bundles}. Sheaves are mathematical abstractions of how data over different parts of a continuous base space are glued together \cite{ghristElementaryAppliedTopology2014, ghri}, whether the data is distributed, streaming, or in disk; therefore sheaves are a way to express the continuity that the artist preserves. 
We assume that $\mathcal{E}$ is defined in terms of a fixed base and fiber space, for example a class of images or tables with the same schema. The graphic category $\mathcal{\gtotal}$ consists of the set of graphics that an artist \vartist\ is allowed to generate from a given $\mathcal{E}$, which is discussed further in \autoref{sec:artist:construction}.

@@

\begin{definition} The category of open sets $\mathcal{\dbase}$ consists of
  \label{def:category:K} 
  \begin{itemize}
    \item \textit{objects} all open sets $U_{i} \in \{U\}$ in the topological space $(\dbase, \mathscr{T})$), including the empty set $\varnothing$ and the maximum set \dbase. 
    \item \textit{morphisms} inclusion maps from a subspace to a larger space $\iota: U_{i} \hookrightarrow U_{j}$, which is equivalent to $U_{i} \subseteq U_{j}$ and  
  \end{itemize} 
\end{definition}

The inclusion maps $\iota$ can be used to piece together the subsets $U_{i}$, particularly individual points $\dbasepoint \in \dbase$. This is how we provide knowledge of continuity to the artist \vartist\ in a way where the artist does not need all the data at once. This is what allows the artist to generate graphics where the subset of data on view is dynamically updated, such as pan and zoom navigation\cite{NekrasovskiEvaluationPanZoom2006} and sliding windows on streaming data\cite{crouchDynamicGraphsSlidingwindow2013,chuTimeSeriesSegmentation1995}. \note{does this actually go down in the composition section?}

@@

\begin{definition} The category of sets of sections $\Gamma(U_i,\dtotal)$, where $U_i$ are objects in $\mathcal{\dfiber}$, form a subcatagory of the category of sets \textbf{Set}. The morphisms between objects of this subcategory are restriction maps $\iota: \Gamma(U_j, \dtotal) \rightarrow \Gamma(U_i, \dtotal)$ where $U_{i} \subseteq U_{j}$.
\end{definition}

As mentioned in \autoref{sec:related-work:fiber-bundles}, the fiber space is embedded in the total space $\dfiber \hookrightarrow \dtotal$. The sections $\Gamma(\dbase, \dtotal)$ are called the global sections of $\dtotal$ \cite{SheafMathematics2021,spanier1989algebraic}. Each section is how we represent a unique dataset, where the sets of sections $\Gamma$ is the space of all datasets that have the same data fields in the \dfiber\ and continuity as encoded in\dbase\. For example, Spivak describes tables that have the same schema and discrete index as sections of the same fiber bundle\cite{spivakSIMPLICIALDATABASES}.

\begin{definition} The category $\mathcal{\dtotal}$ consists of 
  \begin{itemize}
    \item\textit{object} the sheaf\cite{SheafMathematics2021,ghristElementaryAppliedTopology2014} $\sheaf_{\dtotal}: \mathcal{K}^{op} \rightarrow \Gamma(\mathcal{K}, \dtotal)$
    \item \textit{morphisms} arbitrary endomorphic\cite{fongInvitationAppliedCategory2019} bundle maps $\varphi: \dtotal\rightarrow \dtotal$. When the sheaf is restricted to the stalk\cite{StalkSheaf2019} $\mathscr{F}_{k}$ over a point $\dbasepoint\in \dbase$, then allowable morphisms are $\varphi \in Hom_{\mathcal{\dfiber}}(\dfiber,\dfiber)$. 
  \end{itemize}
\end{definition}

Arbitrary morphisms on sheafs $\varphi:\mathcal{\dtotal} \rightarrow \mathcal{\dtotal}$ are natural transforms \cite{SheafMathematics2021,bradleyWhatNaturalTransformation}, meaning that the morphisms $\varphi$ compose with the inclusion maps $\iota$ \cite{fongInvitationAppliedCategory2019}. As defined in \autoref{def:category:Tau}, the sets of sections $\Gamma(U_{i},\dtotal)$ are objects of the category $\mathcal{T}$.

@@

\begin{figure}[!h]
  \centering
  \subfloat[\dbasec_a \sqcup_{\dbasec_c} \dase_b]{\includegraphics[width=1.2in]{union_k.png}%
  \label{fig_first_case}}
  %\hfil
  \subfloat[\dbasec_c=\varnothing]{\includegraphics[width=1.2in]{disjoint_k.png}%
  \label{fig_second_case}}
  \caption{\note{probably doesn't need the k=0}}
  \label{fig_sim}
\end{figure}

>>>>

Formally, a fiber bundle is a mathematical structure $(\dtotalc, \dbasec, \pi, \dfiberc)$
\begin{equation}
  \label{eq:fiber_bundle}
  \begin{tikzcd}
      \dfiberc \arrow[r, hook] & \dtotalc \arrow[r, "\pi"] & \dbasec
  \end{tikzcd}
\end{equation}

\begin{definition}
The \textcolor{base}{base space} \dbasec\ is a topological space, which means it is a set $\dbase$ with a collection of open sets \cite{TopologyVsTopology} surrounding each element in the set. Open sets are a collection of subsets in a set $X$, including the empty set, the whole set $X$, and every union and intersection of its member subsets. For each point $\dbasepoint in \dbase$, there is a collection of subsets $\{U\} \subset \dbase$ such that $x \in U \subseteq \dbase$ \cite{spanier1989algebraic, bradleyTopologyCategoricalApproach2020}  
\end{definition}
\note{intuition about base space}

@@

\begin{equation}
  \begin{tikzcd}
      \dfiberc \arrow[r, hook] & \dtotalc \arrow[d, "\pi"'] \\
                        & \dbasec \arrow[u, "\dsectionc"', bend right]
  \end{tikzcd}
\end{equation}

How these spaces fit together is illustrated in \autoref{fig:related-work:fiber-bundle}, wherein values lives in the fiber $\dfiber\subseteq\dtotal$. In this example, the fiber \dfiber\ is the cartesian product of two sets $F_{0}\times F_{1}$ where each fiber $F_{i}$ is ....

\subsubsection{Sheaves}
The set of all sections sections $\{\dsection^{i}: \U_{j}\rightarrow \dtotal\}|_{i\in I}$ over an arbitrary open set $U_{j}$ is denoted $\Gamma(U_{\dbasepoint}, \dtotal)$. The set of all sections over the total space \dbase\ is denoted $\Gamma(\dbase, \dtotal)$.

@@

\section{Artist} 
In this section, we use category theory to formally express the implicit assumptions that visualization library components make in transforming data into graphical representations. We propose that the visualization transformation can be modeled as a functor, which we call the \textcolor{artist}{Artist} $\mathcal{\vartist}$\footnote{We call this transformation the artist because the \mintinline{python}{Artist} object in Matplotlib\cite{hunterMatplotlib2DGraphics2007}}. The artist \vartist\ converts data, which we model as the sheaf $\sheaf_{\dtotal}$ to graphics, which we model as the sheaf $\sheaf_{\vartist}$. We formulate this association as

\begin{equation}
  \vartist: \sheaf_{\dtotal} \rightarrow \sheaf_{\vartist}
  \label{eq:artist:artist}
\end{equation}

In this section we describe the structure of the data and graphic spaces, which we formulate as objects in categories and the morphisms between those objects. We then use these definitions to express the structure that an artist must preserve. Finally, we propose that artists can be composed by manufacturing new artists based on various types of input data.

\subsection{Categorical Artist}
\label{sec:artist:category}

@@

\begin{definition} The category of open sets $\mathcal{\dbase}$ consists of
  \label{def:category:K} 
  \begin{itemize}
    \item \textit{objects} all open sets $U_{j} \in \{U\}$ in the topological space $(\dbase, \mathscr{T})$), including the empty set $\varnothing$ and the maximum set \dbase. 
    \item \textit{morphisms} inclusion maps from a subspace to a larger space $\iota: U_{j} \hookrightarrow U_{j}$, which is equivalent to $U_{j} \subseteq U_{j}$ and  
  \end{itemize} 
\end{definition}

The inclusion maps $\iota$ can be used to piece together the subsets $U_{i}$, particularly individual points $\dbasepoint \in \dbase$. This is how we provide knowledge of continuity to the artist \vartist\ in a way where the artist does not need all the data at once. This is what allows the artist to generate graphics where the subset of data on view is dynamically updated, such as pan and zoom navigation\cite{NekrasovskiEvaluationPanZoom2006} and sliding windows on streaming data\cite{crouchDynamicGraphsSlidingwindow2013,chuTimeSeriesSegmentation1995}.

@@

\begin{definition} The category \mathcal{\dtotalc} is the set of sets of sections $\{\Gamma(U_j,\dtotal)\}|U_j \in \mathcal{dbase}$. The category \mathcal{\dtotal} is a subcatagory of the category of sets \textbf{Set}. The morphisms between objects of $\mathcal{\dtotal}$ are restriction maps $\iota: \Gamma(U_j, \dtotal) \rightarrow \Gamma(U_i, \dtotal)$ where $U_{i} \subseteq U_{j}$.
\end{definition}

Each section is how we represent a unique dataset, where the sets of sections $\Gamma$ is the space of all datasets that have the same data fields in the \dfiber\ and continuity as encoded in\dbase\. For example, Spivak describes tables that have the same schema and discrete index as sections of the same fiber bundle\cite{spivakSIMPLICIALDATABASES}.

Arbitrary morphisms on sheafs are natural transforms \cite{SheafMathematics2021,bradleyWhatNaturalTransformation}, meaning that the morphisms $\varphi$ compose with the inclusion maps $\iota$ \cite{fongInvitationAppliedCategory2019}. As defined in \autoref{def:category:Tau}, the sets of sections $\Gamma(U_{i},\dtotal)$ are objects of the category $\mathcal{T}$.

@@

\begin{figure}[!h]
  \centering
  \subfloat[\dbasec_a \sqcup_{\dbasec_c} \dbase_b]{\includegraphics[width=1.2in]{union_k.png}%
  \label{fig_first_case}}
  %\hfil
  \subfloat[\dbasec_c=\varnothing]{\includegraphics[width=1.2in]{disjoint_k.png}%
  \label{fig_second_case}}
  \caption{\note{probably doesn't need the k=0}}
  \label{fig_sim}
\end{figure}
